Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-13 paper/sec-analysis.tex
work on the settings section

In this section we motivate the various
kinematical cuts for the various categories.
%
As mentioned before, the cuts in each category have been separately
optimised in an inclusive analysis: it is only once we have determined
which is the dominant category that the analysis is made exclusive.
%
In this work, the results for each category have been derived
from the inclusive analysis,
and we show the various kinematic distributions immediately
below the corresponding cut has been applied.

In this section we only show a representative subset of
all available kinematic distributions, the complete
set of comparison plots can be found in the following
webpage:
\begin{center}
\url{http://www2.physics.ox.ac.uk/HH4b}
  \end{center}
In this webpage we also show the distribution of all
kinematic variables, including jet substructure variables,
which we don't cut explicitly but that are provided as
input to the multivariate analysis.
%
Also, we show here results for the boosted and resolved categories,
a comparison which allows to understand the cuts used for the
intermediate category as well.

We will compare kinematic variables at each step of the analysis
for the boosted and for the resolved categories.
%
The corresponding plots for the intermediate category can be found in the webpage
mentioned above.
%
In the following comparisons, we compare related
kinematical distributions in the two categories.

First of all,
in Fig.~\ref{fig:cutplots1} we show
the $p_T$ distribution of the 
of the
  leading and subleading large-$R$ jet in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  As we will do in all this section,
  all distributions have been normalized to their total integral.
  %
  We see that the background distribution
  falls off rather steeply than the signal, therefore
  suggesting that the boosted region should be
  advantageous to suppress the background.
  %
  This is true both for the leading and the subleading
  large-$R$ jets.
  %
  These distributions justify our choice of $p_T \ge 200$ GeV
  for the large-$R$ jets in the boosted category.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_H0_boost_C1.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_H1_boost_C1.pdf}
\caption{\small  The $p_T$ distribution of the
  leading (left plot) and
  subleading (right plot) large-$R$ jet in the boosted category.
  %
  We comparing
  the shapes of the distribution in signal and background events,
  where both
  distributions have been normalized to their total integral.
}
\label{fig:cutplots1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

In Fig.~\ref{fig:cutplots1} we also show the distribution
of the invariant mass of the leading Higgs candidates.
%
The signal is of course well peaked at $m_h=125$ GeV, the physical
mass, while the background does not display any structure.
%
Note that the smearing, which mimics detector resolution effects,
leads to a rather broad distribution which reduces the usefulness
of the Higgs mass window cut.
%
However, the information of the different shape of the $m_{h}$
distribution will still be exploited by the MVA.

As explained above, in the boosted category we also require the two
leading subjets of the large-$R$ jet to be relatively
hard, in particular that should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig ... we show the distribution in $p_T$ of the leading
and subleading sub-jets in the leading large-$R$ jet in events
of the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background,
motivating our cut.

{\bf missing plots: pt subjet distribution}

Let's now consider the resolved category.
%
To begin with, here it is important to understand the $p_T$ distribution
of the four leading small-$R$ jets of the event.
%
Given that in general the boost from the Higgs decays is moderate,
the two subleading jets might be not too hard, and thus it
should be important to ensure that our $p_T^{\rm min}$ cut
is not too strong.
%
In Fig. we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events.
%
We see that the third and four leading jets are relatively soft.
%
Also, it is useful to show the distribution in rapidity of the
individual jets

{\bf missing plots: pt jet distribution, rapidity distribution}

Next in Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading Higgs candidates, before the Higgs mass window
cut is applied, for the resolved and boosted categories.

@@

Another important kinematical distribution of this process is the invariant mass
of the di-Higgs system.
%
First of all because this is a direct measure of the boost of the system,
and also because in many BSM scenarios this distribution can be substantially
modified as compared to the SM case.
%
Next, we show in
Fig.~\ref{fig:mhh} the invariant mass distribution of the Higgs pairs,
comparing the resolved category (left) with the boosted category (right)
%
This is interesting since it determines the degree of boost that the two
Higgs candidates will have after non-resonant production.
%
In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as for the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this with the MVA in the next section.
%
For the boosted category the trend of the $m_{hh}$ distribution
is very different because of the selection, with the
distribution now peaking at much higher values.
%
In this case signal and background distributions
look reasonably similar.
%
Note that the $m_{HH}$ distribution as a kinematical
cut-off at $m_{HH}^{\rm min}=250$ GeV.

@@

>>>>

In this section we motivate the various
kinematical cuts applied to the different categories.
%
These cuts have been separately
optimised in an inclusive analysis, since it is only
once we have determined
the dominant category that the analysis is made exclusive.
%
Here we show the various kinematic distributions immediately
below the corresponding cut has been applied.
%
Only a representative subset of
all available kinematic distributions will be shown here, the complete
set of plots, including the distribution of all substructure
variables used in the MVA later, have been collected
in this webpage:
\begin{center}
\url{http://www2.physics.ox.ac.uk/HH4b}
  \end{center}
In this section we restrict to the discussion of the
kinematic distributions of the boosted
and resolved categories, the corresponding results for the
intermediate category are collected in the same webpage.

First of all,
in Fig.~\ref{fig:cutplots1} we show
the $p_T$ distribution of the 
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  Distributions have been normalized to their total integral.
  %
  We see that the background distribution
  falls off rather steeply than the signal at large $p_T$
  values, therefore
  providing evidence that exploiting the boosted region will be
  advantageous to enhance the signal over background ratio.
  %
  This is true both for the leading and the subleading
  large-$R$ jets.
  %
  These distributions justify our choice of $p_T \ge 200$ GeV
  for the large-$R$ jets in the boosted category.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_H0_boost_C1.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_H1_boost_C1.pdf}
\caption{\small  The $p_T$ distribution of the
  leading (left plot) and
  subleading (right plot) large-$R$ jets in the boosted category.
  %
  We comparing
  the shapes of the distribution in signal and background events,
  where both
  distributions have been normalized to their total integral.
}
\label{fig:cutplots1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

{\bf Add missing plots here: pt of subjets(boosted), pt of
  small-R jets (resolved), rapidity of the large-$R$ jets,
rapidity of the small-$R$ jets.}

In the boosted category we require the two
leading subjets of the large-$R$ jet to be relatively
hard, in particular that should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in FigAAA
we show the distribution in $p_T$ of the leading
and subleading sub-jets in the leading large-$R$ jet in events
of the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background,
motivating our choice of subjet $p_T$ cut.

In the intermediate category,
it is important to understand the $p_T$ distribution
of the four leading small-$R$ jets of the event.
%
Given that in general the boost from the Higgs decays is moderate,
the two subleading jets might be not too hard, and thus it
should be important to ensure that our $p_T^{\rm min}$ cut
is not too strong.
%
This also has important implications for the experimental
trigger requirements in this category.
%
In Fig. we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events.
%
We see that the third and four leading jets are relatively soft.
%
Therefore, we can conclude that....

Another important distribution is the rapidity of the large-$R$
(in the boosted category) and the small-$R$ (in the resolved
category) jets.
%
We want to understand by how much the signal efficiency is reduced
by the restriction of the central region, and how would
the signal yield be enhanced if tracking could be extended
to forward rapidities in ATLAS and CMS.
%
Also, it is useful to show the distribution in rapidity of the
individual jets in the resolved category.
%
This comparison is shown in FigAAA.

Next in Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut is applied, for the resolved and boosted categories.
%
The signal distribution is of course peaked at the
nominal Higgs mass of $m_h=125$ GeV.
%
In the boosted category, the background shows no particular
structure, while remarkable in the resolved category the background
also shows a peak-like structure, artificially induced by the
acceptance kinematical cuts on the jets.
%
Note that the applied smearing, which mimics detector resolution effects,
leads to a rather broad distribution which reduces the usefulness
of the Higgs mass window cut.
%
While our Higgs mass window cut is relatively loose,
information of the different shape of the $m_{h}$
distribution will still be exploited by the MVA.
%
In Sect.~\ref{sec:optimisation} we discuss how the results
of the analysis is modified if a more aggressive
detector resolution is assumed (corresponding
to a smaller value of the smearing).

@@

Another important kinematical distribution of this process is the invariant mass
of the di-Higgs system.
%
First of all because this is a direct measure of the boost of the system,
and also because in many BSM scenarios this distribution can be substantially
modified as compared to the SM case, for example in the presence
of certain dimension-6 EFT operators~\cite{Azatov:2015oxa}.
%
Indeed, one important advantage of the $4b$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching rations,
such as $2b2\gamma$.
%
With this motivation, we show in
Fig.~\ref{fig:mhh} the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved category (left) with the boosted category (right).

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as for the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this with the help of the MVA.
%
For the boosted category the trend of the $m_{hh}$ distribution
is different because of the jet selection cuts, with the
distribution now peaking at higher values.
%
In this case signal and background distributions
look reasonably similar.
%
Note that at parton-level the $m_{hh}$ distribution as a kinematical
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is modified by both
shower effects and by detector resolution effects.
%
While we don't cut in this quantity, it is
one of the inputs for the MVA to help to improve signal discrimination.

@@

Another interesting distribution is the transverse momentum of
the di-Higgs system.
%
In Fig.~\ref{fig:pthh} we show the $p_T^{hh}$
distribution
for the resolved and boosted categories.
%
Note that in our LO simulation this distribution is generated purely
by the parton shower - a more refined calculation would require
either the matching with higher-multiplicity matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta}, to treat properly the first hard emission.
%
Again we see that the background has a steeper $p_T^{hh}$ distribution
that the signal, in both categories, thus this variable
should provide some additional discrimination power, and therefore
we will use it as another of the MVA inputs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.48\textwidth]{plots/pt_HH_res_C1.pdf}
  \includegraphics[width=0.48\textwidth]{plots/pt_HH_boost_C1.pdf}
  \caption{\small Left plot: the invariant mass distribution of the Higgs
    pair candidates, $m_{hh}$, comparing signal and background events,
    in the resolved category.
    %
    Right plot: same for the boosted category.
}
\label{fig:pthh}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-13 paper/sec-intro.tex
work on the settings section

In addition to the gluon-fusion production mode, Higgs pairs
can be also produced in the vector-boson fusion
channel $hhjj$~\cite{Contino:2010mh,Dolan:2013rja,Dolan:2015zja},
the associated production modes $hhW$ and $hhZ$, and also in association
with top quark pairs $hht\bar{t}$~\cite{Englert:2014uqa}.
%
All these channels are extremely challenging due to the small production
rates: at 14 TeV, the inclusive total cross-sections are
2 fb for VBF $hhjj$~\cite{Liu-Sheng:2014gxa},
0.5 fb for $W(Z)hh$~\cite{baglio}
and 1.0 for $hht\bar{t}$~\cite{Englert:2014uqa}.
%
Concentrating on the SM production dynamics of Higgs pair
production in the various channels, several studies
have investigate the interplay of various kinematical
distributions towards the optimisation of the
extraction of $\lambda$ from the measured
cross-sections, as well as the improvement
in the signal modeling taking into account
higher-order corrections and theoretical
uncertainties~\cite{Slawinska:2014vpa,Chen:2014xra,Goertz:2013kp,Frederix:2014hta,Dawson:2015oha,Maltoni:2014eza}
%

>>>>

In addition to the gluon-fusion production mode, Higgs pairs
can be also produced in the vector-boson fusion
channel $hhjj$~\cite{Contino:2010mh,Dolan:2013rja,Dolan:2015zja},
the associated production modes $hhW$ and $hhZ$, and also in association
with top quark pairs $hht\bar{t}$~\cite{Englert:2014uqa}.
%
All these channels are extremely challenging due to the small production
rates: at 14 TeV, the inclusive total cross-sections are
2 fb for VBF $hhjj$~\cite{Liu-Sheng:2014gxa},
0.5 fb for $W(Z)hh$~\cite{baglio}
and 1.0 for $hht\bar{t}$~\cite{Englert:2014uqa}.
%
Concentrating on the SM production dynamics of Higgs pair
production in the various channels, several studies
have investigate the interplay of various kinematical
distributions towards the optimisation of the
extraction of $\lambda$ from the measured
cross-sections, as well as the improvement
in the signal modeling taking into account
higher-order corrections and theoretical
uncertainties~\cite{Slawinska:2014vpa,Chen:2014xra,Goertz:2013kp,Frederix:2014hta,Dawson:2015oha,Maltoni:2014eza,Maierhofer:2013sha}
%
